---
title: "Urology"
author: "Jacqueline Grout"
echo: false
warning: false
message: false
format: html
editor: visual
---

```{r}
#| echo: false
library(sf)
library(tmap)
library(dplyr)
library(stringr)
library(readxl)
library(gt)
library(janitor)
library(readr)
library(tidyr)
library(purrr)
library(tidyverse)
library(ggrepel)

load(file = "data/projectimage.RData")
source("R/functions.R")

#midlandsoutpatients <- readRDS("midlandsoutpatients.rds")
```

## Executive Summary

Yet to be written ...

### Key Findings

Yet to be written ...

### Implications/Recommendations

Yet to be written ...

## Introduction

There are concerns that the provision of some urology services, particularly functional urology,  is not meeting the needs of patients adequately. This gap in provision could lead to delayed or prevented diagnosis and treatment resulting in poor quality of life and negative health outcomes for patients. The aim of this project is to produce a quantitative analysis to explore and quantify the underlying level of need in the Midlands for a selection of common functional urological conditions and using that to begin to quantify the scale of some of the gaps in Urology provision.

## Methodology

### Underlying need

Prevalence studies from the literature for the selected conditions were considered with the aim to obtain detailed age and gender specific prevalence rate estimates.

Benign Prostatic Hyperplasia:

Prevalence rates were taken from, "The global burden of lower urinary tract symptoms suggestive of benign prostatic hyperplasia: A systematic review and meta-analysis" by Lee, Chan & Lai (2017).

<https://www.nature.com/articles/s41598-017-06628-8>

This article was chosen as it provided prevalence rates consistent with other literature, but included prevalence rates by age. This condition is only relevant to males.

```{r}

select(prev_bph, -prevpercent_female) |>
  gt()|>
  cols_label(age_group_bph = "Age Group",
             prevpercent_male = "Prevalence (%)") |>
  tab_header(
    title = 'Benign Prostatic Hyperplasia (BPH) Prevalence'
  ) |>
  tab_source_note(source_note = "Source: Lee, Chan & Lai (2017)")
```

Overactive Bladder:

Prevalence rates were taken from, "Prevalence of the Overactive Bladder Syndrome by Applying the International Continence Society Definition" by Temml, Heidler, Ponholzer and Madersbacher (2005).

<https://ginegap.com/materiales-cursos/bibliografia/m1-c1/52_Temml_2005.pdf>

This article was chosen as it provided prevalence rates consistent with other literature, used the International Continence Society (ICS) definitions, and included prevalence rates by both age and gender.

```{r}
prev_oab |>
  gt()|>
  cols_label(age_group_oab = "Age Group",
             prevpercent_male = "Male Prevalence (%)",
             prevpercent_female = "Female Prevalence (%)")|>
  tab_header(
    title = 'Overactive Bladder Prevalence'
  ) |>
  tab_source_note(source_note = "Source: Temml, Heidler, Ponholzer and Madersbacher (2005)")
```

Post prostatectomy incontinence:

Prevalence rates were taken from, "Prevalence analysis of urinary incontinence after radical prostatectomy and influential preoperative factors in a single institution" by Robles, Hevia, Algarra, Diez-Cabellero and Pascual (2017).

<https://www.tandfonline.com/doi/full/10.1080/13685538.2017.1369944>

This article was chosen as it also used the ICS definitions. This condition relates only to male patients, of whom the article states that 23% remain incontinent 12 months after a prostatectomy.

For both Benign Prostatic Hyperplasia (BPH) and Overactive Bladder the prevalence rates obtained from the literature were applied to the GP practice list sizes to obtain prevalence estimates for each practice, taking account of the gender and age profile of the practice.

Applying the rates to the GP list sizes this gives an indication of the underlying level of need for each GP practice. The population need was then aggregated to generate need for each sub-ICB and each ICB with the intention of providing quantification of underlying need at both strategic and a more local/operational level.

### Treatment to need ratios

Treatment to need ratios have been calculated by directly comparing levels of condition specific treatment activity to the prevalence estimates. The ratios are activity divided by need.

The analysis has considered hospital surgical treatments (including inpatient, day case and outpatient treatments) using surgical procedure data as well as primary care management using condition specific prescribing data.

For the Benign Prostatic Hyperplasia and Overactive Bladder conditions the surgical procedure activity was a count by GP practice of the number of patients with at least one spell (Elective or Daycase) for a relevant procedure in the 5 year period (April 2018 to March 2023) with a Urology (101) treatment function code and a relevant diagnosis code (see appendix). Where a patient had more than one spell then the most recent spell was used to assign them to a GP practice.

For the Post Prostatectomy Incontinence condition the surgical procedure activity was one of two definitions; 1) a count by GP practice of the number of patients in a 5-year period (April 2013 to March 2023), with at least one spell (Elective or Daycase) in which a procedure took place, with a Urology (101) treatment function code and a relevant diagnosis code (see appendix) in any position, who had previously (1 to 6 years prior, April 2012 to March 2017) had a spell in which they had a prostatectomy procedure; and 2) a count by GP practice of the number of patients in a 5-year period (April 2013 to March 2023), with at least one spell (Elective or Daycase) with a Urology (101) treatment function code and a relevant primary diagnosis code (see appendix), who had previously (1 to 6 years prior, April 2012 to March 2017) had a spell in which they had a prostatectomy procedure.

Outpatient Data

The Urology outpatient data rarely includes diagnosis codes (only 4% of attendances). Procedures that take place in outpatients are also not consistently recorded. Consequently it's not possible to give a robust picture of activity in relation to these specific conditions for use in calculating a treatment to need ratio. It is however possible to provide a description of the distribution of Urology outpatient activity across the Midlands region. Outpatient data was obtained from the SUS data as counts of outpatient attendances between April 2018 and March 2023 with a treatment function code of 101.

Prescribing data

Condition specific prescribing data was used as a proxy for / to reflect primary care management of the conditions. Prescribing data was obtained from *OpenPrescribing.net, Bennett Institute for Applied Data Science, University of Oxford, 2023*

Built-in APIs were used from

<https://openprescribing.net/>

This is anoymised data about the drugs prescribed by GPs and drugs are organised and coded into relevant chapters. The detail of the drugs, codes and chapters used for each condition can be seen in the appendix.

The activity was taken as a count of prescription items by sub-ICB over the most recent 5 year period available.

### Analysis by GP practice

During the five-year period of time of the activity data there have been changes in GP practices and their sizes, including closures, mergers etc. When reporting data at GP practice level, any practice with a list size that has changed by more than 20% over this 5 year period has been excluded from the reporting. The data has however been included in reporting at sub-ICB and ICB level.

## FINDINGS:

### Inpatients

#### Benign Prostatic Hyperplasia:

##### Prevalence

The prevalence for each sub-ICB was calculated by dividing the age and gender specific prevalence calculated for the sub-ICB and dividing this by the overall total list size for the respective sub-ICB. The highest overall prevalence rates are in Stafford and Surrounds (05V) sub-ICB and Herefordshire and Worcestershire (18C) ICB. Prevalence rates are the lowest in Leicester City (04C) sub-ICB.

```{r}
#| echo: false

#subicb_bph <- readRDS("subicb_bph.rds")

tmap_options(check.and.fix = TRUE)
tmap_mode("view") #non-interactive maps = "plot"

bphprevmap <- tm_shape(subicb_bph) +
tm_fill("overall_prev",title="Prevalence BPH",style = "pretty", palette = "Blues") +
tm_borders("grey25", alpha=.5) +
tm_basemap(server = c('OpenStreetMap'))

bphprevmap
```

```{r}
select(subicb_bph, SICBL22NM, bphprevnumtotal, list_size_bph_total, overall_prev) |>
  st_drop_geometry() |>
  arrange(SICBL22NM) |>
  gt() |>
    cols_label(
    SICBL22NM = md("Sub-ICB"),
    bphprevnumtotal = md("BPH number"),
    list_size_bph_total = md("GP Practice List Size"),
    overall_prev = md("Prevalence (%)")
  ) |>
  cols_align(
    align = "right",
    columns = 2:4
  ) |>
  fmt_number(columns = 2:3,
             decimals=0) |> 
  fmt_number(columns = 4:4,
             decimals=2) |>
tab_header(
    title = 'Benign Prostatic Hyperplasia - Prevalence'
  ) |>
  tab_source_note(source_note = "Data: SUS & Patients Registered at a GP Practice (NHS Digital)")
```

##### Activity to Need Ratio

The BPH activity to need ratios indicate that Bassetlaw (02Q) sub-ICB and Herefordshire and Worcestershire (18C) ICB have the highest level of activity relative to the underlying need. The lowest level of activity relative to the underlying need is in Leicester City (04C) sub-ICB and Stoke-on-Trent (05W) sub-ICB.

```{r}
#| echo: false
library(StrategyUnitTheme)

#tmap_mode("plot")

tm_shape(subicb_bph_ratio) + 
tm_fill("activityneedratio",title="Activity/Need Ratio BPH",style = "pretty", palette = "-viridis") +
tm_borders("grey25", alpha=.5) +  
tm_shape(bph_providers_sf) + 
tm_dots(
          #col = "grey", size = 0.05,legend.show = T)+
          col= "short_role",size = 0.05, title = "Providers") +
tm_basemap(server = c('OpenStreetMap'))
```

Funnel plots

```{r}


data <-ICBTOBPH |>
  filter(reg22nm =="Midlands") |>
  mutate(activityneedratio = num_bph_spells / bphprevnumtotal)
         
ubar <- sum(data$num_bph_spells)/sum(data$bphprevnumtotal)


#sub_icb <- midlands_sub_icbs[[i]]

cl_fn <- function(ubar, limit) {
  \(x) ubar + limit * sqrt(ubar / x)
}


plots <-midlands_sub_icbs |>
  set_names() |>
  map(\(sub_icb) {
    ggplot(data, aes(bphprevnumtotal, activityneedratio)) +
      geom_jitter(
        data = \(.x) filter(.x, sub_icb_location_code == sub_icb),
        fill = "#f9bf07",
        colour = "black",
        shape="circle filled"
      ) +
      # geom_jitter(
      #   data = \(.x) {
      #     .x |>
      #       filter(sub_icb_location_code != sub_icb)
      #   },
      #   alpha = 0.1,
      #   size = 0.5
      #     ) +
      geom_hline(aes(yintercept=ubar),size=1) +
      geom_text_repel(
        data = \(.x) {
          .x |>
            filter(
              sub_icb_location_code == sub_icb,
              FALSE |
                activityneedratio > cl_fn(ubar, 3.0)(bphprevnumtotal) |
                activityneedratio < cl_fn(ubar, -3.0)(bphprevnumtotal)
            )
        },
        aes(label = org_code)
      ) +
      geom_function(fun = cl_fn(ubar, +1.96), size = 0.5, linetype=5) +
      geom_function(fun = cl_fn(ubar, -1.96), size = 0.5, linetype=5) +
      geom_function(fun = cl_fn(ubar, +3.00), size = 0.5, linetype="solid") +
      geom_function(fun = cl_fn(ubar, -3.00), size = 0.5, linetype="solid") +
      xlab("Prevalence Number") +
      ylab("Activity to Need Ratio")+
      ylim(-0.1, 0.1)+
      labs(title = "Funnel plot for Benign Prostatic Hyperplasia", 
           subtitle = paste0("Sub-ICB: ",sub_icb)) +
      su_theme() 
  })

purrr::walk(plots, plot)
```

##### Activity Gap

The variation in activity to need ratios results in an activity gap and in order to attempt to quantify this activity gap, the activity to need ratio for the sub-ICB with the highest ratio has been applied to all other sub-ICBs and an estimated number of inpatient spells calculated as well as the % increase in activity that this represents. The table below shows that Leicester City (04C) sub-ICB would need a 119% increase in activity to increase its activity to need ratio to the same as Bassetlaw (02Q) sub-ICB.

```{r}
sub_icbs_mid <- select(ICBTOBPH,sub_icb_location_code, new_sub_icb_location_name, icb22, icb22nm) |>
  filter(sub_icb_location_code %in% midlands_sub_icbs)|>
  filter(icb22 != "NA")|>
  unique()

bph_activity_gap <- bph_activity |> 
  filter(sub_icb_location_code %in% midlands_sub_icbs) |>
  mutate(uq_activity_gap = (quantile(activityneedratio,prob=0.75,type=1)*bphprev)-bph_spells,
         uq_gap_percentage_change = (uq_activity_gap/bph_spells)*100,
         activity_gap = (max(activityneedratio)*bphprev)-bph_spells,
         gap_percentage_change = (activity_gap/bph_spells)*100)|>
  left_join(sub_icbs_mid) |>
  select(new_sub_icb_location_name,bphprev,list_size_bph,bph_spells,activityneedratio,activity_gap,gap_percentage_change)
#,uq_activity_gap, uq_gap_percentage_change

bar_yields <- bph_activity_gap |> 
  rowwise() |>
  ungroup() |> 
  mutate(
    bar = gap_percentage_change,
    color = col_pal(bar,"ggsci::blue_material"),
    bar_chart = bar_chart(bar, color = color),
    bar_chart = map(bar_chart, ~gt::html(as.character(.x)))
  ) |>
  select(-bar, -color)

bph_activity_gap_table <- 
  bar_yields |>
  gt()|>
  cols_width(c(bar_chart) ~ px(100),
             c(bphprev) ~ px(50),
             c(list_size_bph) ~ px(50),
             c(bph_spells) ~ px(50),
             c(activityneedratio) ~ px(50),
             c(activity_gap) ~ px(50),
             c(gap_percentage_change) ~ px(50)
  ) %>% 
  fmt_number(columns = 2:4,
             decimals=0) |>
  fmt_number(columns = 5:8,
             decimals=2) |>
  cols_label(
    new_sub_icb_location_name=md("Sub-ICB"),
    bphprev = md("Prevalence"),
    list_size_bph = md("List Size"),
    bph_spells = md("Spells"),
    activityneedratio = md("Activity to Need ratio"),
    activity_gap = md("Activity Gap"),
    gap_percentage_change = md("% change"),
    bar_chart = ""
  ) |>
  cols_align(
    align = "right",
    columns = 2:7
  ) |>
  cols_align(
    align = "left",
    columns = c(bar_chart)
  ) |>
  tab_header(
    title = 'Benign Prostatic Hyperplasia Activity Gap - Midlands'
  ) |>
  tab_source_note(source_note = "Data: SUS??")

bph_activity_gap_table

```

#### Overactive Bladder:

For the Overactive Bladder condition, the highest overall prevalence rates are in North Staffordshire (05G) sub-ICB and Lincolnshire (71E) ICB. Prevalence rates are the lowest in Leicester City 04C.

##### Prevalence

```{r}
#| echo: false

#subicb_oab <- readRDS("subicb_oab.rds")


tm_shape(subicb_oab) +
tm_fill("overall_prev",title="Prevalence OAB",style = "pretty", palette = "Greens") +
tm_borders("grey25", alpha=.5) +
tm_basemap(server = c('OpenStreetMap'))
```

```{r}
select(subicb_oab, SICBL22NM, oabprevnumtotal, list_size_oab_total, overall_prev) |>
  st_drop_geometry() |>
  arrange(SICBL22NM) |>
  gt() |>
    cols_label(
    SICBL22NM = md("Sub-ICB"),
    oabprevnumtotal = md("OAB number"),
    list_size_oab_total = md("GP Practice List Size"),
    overall_prev = md("Prevalence (%)")
  ) |>
  cols_align(
    align = "right",
    columns = 2:4
  ) |>
  fmt_number(columns = 2:3,
             decimals=0) |> 
  fmt_number(columns = 4:4,
             decimals=2) |>
tab_header(
    title = 'Overactive Bladder - Prevalence'
  ) |>
  tab_source_note(source_note = "Data: SUS & Patients Registered at a GP Practice (NHS Digital)")
```

Where conditions relate to both male and female patients the results will be stratified by gender to inform an assessment of potential gender inequity in treatment access/provision.

##### Activity to Need Ratio

The overactive bladder activity to need ratios indicate that Bassetlaw (02Q) sub-ICB and East Leicestershire and Rutland (03W) sub-ICB have the highest level of activity relative to the underlying need. The lowest level of activity relative to the underlying level of need is in Leicester City (04C) sub-ICB and Birmingham and Solihull (15E) ICB.

```{r}
#| echo: false

#subicb_oab_ratio <- readRDS("subicb_oab_ratio.rds")
#oab_providers_sf <- readRDS("oab_providers_sf.rds")

tm_shape(subicb_oab_ratio) +
tm_fill("activityneedratio",title="Activity/Need Ratio OAB",style = "pretty", palette = "-viridis") +
tm_borders("grey25", alpha=.5) +
tm_shape(oab_providers_sf) + 
tm_dots(col = "short_role",size = 0.05, title = "Providers")+
tm_basemap(server = c('OpenStreetMap'))
```

Funnel Plots

```{r}
data <-ICBTOOAB |>
  filter(reg22nm =="Midlands") |>
  mutate(activityneedratio = num_oab_spells / oabprevnumtotal)
         
ubar <- sum(data$num_oab_spells)/sum(data$oabprevnumtotal)


#sub_icb <- midlands_sub_icbs[[i]]

cl_fn <- function(ubar, limit) {
  \(x) ubar + limit * sqrt(ubar / x)
}


plots <-midlands_sub_icbs |>
  set_names() |>
  map(\(sub_icb) {
    ggplot(data, aes(oabprevnumtotal, activityneedratio)) +
      geom_jitter(
        data = \(.x) filter(.x, sub_icb_location_code == sub_icb),
        fill = "#f9bf07",
        colour = "black",
        shape="circle filled"
      ) +
      # geom_jitter(
      #   data = \(.x) {
      #     .x |>
      #       filter(sub_icb_location_code != sub_icb)
      #   },
      #   alpha = 0.1,
      #   size = 0.5
      #     ) +
      geom_hline(aes(yintercept=ubar),size=1) +
      geom_text_repel(
        data = \(.x) {
          .x |>
            filter(
              sub_icb_location_code == sub_icb,
              FALSE |
                activityneedratio > cl_fn(ubar, 3.0)(oabprevnumtotal) |
                activityneedratio < cl_fn(ubar, -3.0)(oabprevnumtotal)
            )
        },
        aes(label = org_code)
      ) +
      geom_function(fun = cl_fn(ubar, +1.96), size = 0.5, linetype=5) +
      geom_function(fun = cl_fn(ubar, -1.96), size = 0.5, linetype=5) +
      geom_function(fun = cl_fn(ubar, +3.00), size = 0.5, linetype="solid") +
      geom_function(fun = cl_fn(ubar, -3.00), size = 0.5, linetype="solid") +
      xlab("Prevalence Number") +
      ylab("Activity to Need Ratio")+
      ylim(-0.025, 0.025)+
      labs(title = "Funnel plot for Overactive Bladder", 
           subtitle = paste0("Sub-ICB: ",sub_icb)) +
      su_theme() 
  })

purrr::walk(plots, plot)
```

##### Activity Gap

As with Benign Prostatetic Hyperplasia there is an activity gap for Overactive Bladder and in order to attempt to quantify this activity gap, the activity to need ratio for the sub-ICB with the highest ratio has been applied to all other sub-ICBs and an estimated number of inpatient spells calculated as well as the % increase in activity that this represents. The table below shows that Leicester City (04C) sub-ICB would need a 163% increase in activity to increase its activity to need ratio to the same as Bassetlaw (02Q) sub-ICB.

```{r}
sub_icbs_mid <- select(ICBTOOAB,sub_icb_location_code, new_sub_icb_location_name, icb22, icb22nm) |>
  filter(sub_icb_location_code %in% midlands_sub_icbs)|>
  filter(icb22 != "NA")|>
  unique()

oab_activity_gap <- oab_activity |> 
  filter(sub_icb_location_code %in% midlands_sub_icbs) |>
  mutate(uq_activity_gap = (quantile(activityneedratio,prob=0.75,type=1)*oabprev)-oab_spells,
         uq_gap_percentage_change = (uq_activity_gap/oab_spells)*100,
         activity_gap = (max(activityneedratio)*oabprev)-oab_spells,
         gap_percentage_change = (activity_gap/oab_spells)*100)|>
  left_join(sub_icbs_mid) |>
  select(new_sub_icb_location_name,oabprev,list_size_oab,oab_spells,activityneedratio,activity_gap,gap_percentage_change)
#,uq_activity_gap, uq_gap_percentage_change

bar_yields <- oab_activity_gap |> 
  rowwise() |>
  ungroup() |> 
  mutate(
    bar = gap_percentage_change,
    color = col_pal(bar,"ggsci::green_material"),
    bar_chart = bar_chart(bar, color = color),
    bar_chart = map(bar_chart, ~gt::html(as.character(.x)))
  ) |>
  select(-bar, -color)

oab_activity_gap_table <- 
  bar_yields |>
  gt()|>
  cols_width(c(bar_chart) ~ px(100),
             c(oabprev) ~ px(50),
             c(list_size_oab) ~ px(50),
             c(oab_spells) ~ px(50),
             c(activityneedratio) ~ px(50),
             c(activity_gap) ~ px(50),
             c(gap_percentage_change) ~ px(50)
  ) %>% 
  fmt_number(columns = 2:4,
             decimals=0) |>
  fmt_number(columns = 5:8,
             decimals=2) |>
  cols_label(
    new_sub_icb_location_name=md("Sub-ICB"),
    oabprev = md("Prevalence"),
    list_size_oab = md("List Size"),
    oab_spells = md("Spells"),
    activityneedratio = md("Activity to Need ratio"),
    activity_gap = md("Activity Gap"),
    gap_percentage_change = md("% change"),
    bar_chart = ""
  ) |>
  cols_align(
    align = "right",
    columns = 2:7
  ) |>
  cols_align(
    align = "left",
    columns = c(bar_chart)
  ) |>
  tab_header(
    title = 'Overactive Bladder Activity Gap - Midlands'
  ) |>
  tab_source_note(source_note = "Data: SUS & Patients Registered at a GP Practice (NHS Digital)")

oab_activity_gap_table
```

#### Post Prostatectomy Incontinence:

##### Prevalence

The post-prostatectomy incontinence prevalence for each sub-ICB was calculated differently to the other conditions, by dividing the age and gender specific prevalence calculated for the sub-ICB and dividing this by the overall total list size for the respective sub-ICB. The highest overall prevalence rates are in Herefordshire and Worcestershire (18C) ICB and Bassetlaw (02Q) sub-ICB. Prevalence rates are the lowest in Leicester City (04C) sub-ICB.

```{r}
#| echo: false

#subicb_ppi <- readRDS("subicb_ppi.rds")

 subicb_ppi <- subicb_ppi |>
   mutate(prev_per_list_size = (overall_prev/list_size_ppi)*1000) 

tm_shape(subicb_ppi) +
tm_fill("prev_per_list_size",title="Prevalence PPI per 1000",style = "pretty", palette = "Reds") +
tm_borders("grey25", alpha=.5) +
tm_basemap(server = c('OpenStreetMap'))
```

```{r}
select(subicb_ppi,SICBL22CD,SICBL22NM,overall_prev,list_size_ppi,prev_per_list_size)|>
  st_drop_geometry() |>
  arrange(SICBL22NM) |>
  select(-SICBL22CD)|>
  gt() |>
  cols_label(
    SICBL22NM = md("Sub-ICB"),
    overall_prev = md("Prostatectomy Prevalence Estimate"),
    list_size_ppi = md("GP List Size"),
    prev_per_list_size = md("Prevalence (per 1000 List Size)")
  ) |>
  cols_align(
    align = "right",
    columns = 2:4
  ) |>
  fmt_number(columns = 2:3,
             decimals=0) |>
  fmt_number(columns = 4:4,
             decimals=2) |>
  tab_header(
    title = 'Post Prostatectomy Incontinence - Prevalence'
  ) |>
  tab_source_note(source_note = "Data: SUS & Patients Registered at a GP Practice (NHS Digital)")

```

##### Activity to Need Ratio

The post prostatectomy incontinence activity to need ratios indicate that Stafford and Surrounds (05V) sub-ICB and Bassetlaw (02Q) sub-ICB have the highest level of activity relative to the underlying need. The lowest level of activity is in North Staffordshire (05G) sub-ICB and Lincolnshire (71E) ICB.

```{r}
#| echo: false

#subicb_ppi_ratio <- readRDS("subicb_ppi_ratio.rds")
#ppi_providers_sf <- readRDS("ppi_providers_sf.rds")


tm_shape(subicb_ppi_ratio) +
  tm_fill("activityneedratio",title="Activity/Need Ratio PPI",style = "pretty", palette = "-viridis") +
  tm_borders("grey25", alpha=.5) +
  tm_shape(ppi_providers_sf) + 
  tm_dots(col = "short_role",size = 0.05, title = "Providers")+
  tm_basemap(server = c('OpenStreetMap'))

```

Funnel plots

The sub-ICB plots below show the activity to need ratios for each GP practice plotted according to the underlying need for the condition and with confidence intervals for the whole of the Midlands region. GP practices that are outside the limits for the Midlands region are labelled. Within the two sub-ICBs with the highest activity to need ratios all GP practices are within the confidence limits.

```{r}
data <-ICBTOPPI |>
  filter(reg22nm =="Midlands") |>
  mutate(activityneedratio = num_ppi_spells / numtotal)
         
ubar <- sum(data$num_ppi_spells)/sum(data$numtotal)


#sub_icb <- midlands_sub_icbs[[i]]

cl_fn <- function(ubar, limit) {
  \(x) ubar + limit * sqrt(ubar / x)
}


plots <-midlands_sub_icbs |>
  set_names() |>
  map(\(sub_icb) {
    ggplot(data, aes(numtotal, activityneedratio)) +
      geom_jitter(
        data = \(.x) filter(.x, sub_icb_location_code == sub_icb),
        fill = "#f9bf07",
        colour = "black",
        shape="circle filled"
      ) +
      # geom_jitter(
      #   data = \(.x) {
      #     .x |>
      #       filter(sub_icb_location_code != sub_icb)
      #   },
      #   alpha = 0.1,
      #   size = 0.5
      #     ) +
      geom_hline(aes(yintercept=ubar),size=1) +
      geom_text_repel(
        data = \(.x) {
          .x |>
            filter(
              sub_icb_location_code == sub_icb,
              FALSE |
                activityneedratio > cl_fn(ubar, 3.0)(numtotal) |
                activityneedratio < cl_fn(ubar, -3.0)(numtotal)
            )
        },
        aes(label = org_code)
      ) +
      geom_function(fun = cl_fn(ubar, +1.96), size = 0.5, linetype=5) +
      geom_function(fun = cl_fn(ubar, -1.96), size = 0.5, linetype=5) +
      geom_function(fun = cl_fn(ubar, +3.00), size = 0.5, linetype="solid") +
      geom_function(fun = cl_fn(ubar, -3.00), size = 0.5, linetype="solid") +
      xlab("Prevalence Number") +
      ylab("Activity to Need Ratio")+
      labs(title = "Funnel plot for Post Prostatectomy Incontinence", 
           subtitle = paste0("Sub-ICB: ",sub_icb)) +
      su_theme() 
  })

purrr::walk(plots, plot)
```

Again, as with other conditions, there is an activity gap for Post prostatectomy Incontinence and in order to attempt to quantify this activity gap, the activity to need ratio for the sub-ICB with the highest ratio has been applied to all other sub-ICBs and an estimated number of inpatient spells calculated as well as the % increase in activity that this represents. The table below shows that North Staffordshire (05G) sub-ICB would need a 157% increase in activity to increase its activity to need ratio to the same as Stafford and Surrounds (05V) sub-ICB.

```{r}
sub_icbs_mid <- select(ICBTOPPI,sub_icb_location_code, new_sub_icb_location_name, icb22, icb22nm) |>
  filter(sub_icb_location_code %in% midlands_sub_icbs)|>
  filter(icb22 != "NA")|>
  unique()

ppi_activity_gap <- ppi_activity |> 
  filter(sub_icb_location_code %in% midlands_sub_icbs) |>
  mutate(uq_activity_gap = (quantile(activityneedratio,prob=0.75,type=1)*ppiprevsum)-ppi_spells,
         uq_gap_percentage_change = (uq_activity_gap/ppi_spells)*100,
         activity_gap = (max(activityneedratio)*ppiprevsum)-ppi_spells,
         gap_percentage_change = (activity_gap/ppi_spells)*100)|>
  left_join(sub_icbs_mid) |>
  select(new_sub_icb_location_name,ppiprevsum,ppi_spells,activityneedratio,activity_gap,gap_percentage_change)
#,uq_activity_gap, uq_gap_percentage_change

bar_yields <- ppi_activity_gap |> 
  rowwise() |>
  ungroup() |> 
  mutate(
    bar = gap_percentage_change,
    color = col_pal(bar,"ggsci::red_material"),
    bar_chart = bar_chart(bar, color = color),
    bar_chart = map(bar_chart, ~gt::html(as.character(.x)))
  ) |>
  select(-bar, -color)

ppi_activity_gap_table <- 
  bar_yields |>
  gt()|>
  cols_width(c(bar_chart) ~ px(100),
             c(ppiprevsum) ~ px(50),
             #c(list_size_ppi) ~ px(50),
             c(ppi_spells) ~ px(50),
             c(activityneedratio) ~ px(50),
             c(activity_gap) ~ px(50),
             c(gap_percentage_change) ~ px(50)
  ) %>% 
  fmt_number(columns = 2:3,
             decimals=0) |>
  fmt_number(columns = 4:7,
             decimals=2) |>
  cols_label(
    new_sub_icb_location_name=md("Sub-ICB"),
    ppiprevsum = md("Prevalence"),
  #  list_size_ppi = md("List Size"),
    ppi_spells = md("Spells"),
    activityneedratio = md("Activity to Need ratio"),
    activity_gap = md("Activity Gap"),
    gap_percentage_change = md("% change"),
    bar_chart = ""
  ) |>
  cols_align(
    align = "right",
    columns = 2:6
  ) |>
  cols_align(
    align = "left",
    columns = c(bar_chart)
  ) |>
  tab_header(
    title = 'Post Prostatectomy Incontinence Activity Gap - Midlands'
  ) |>
  tab_source_note(source_note = "Data: SUS & Patients Registered at a GP Practice (NHS Digital)")

ppi_activity_gap_table
```

### Outpatients

There were 2.3 million (2,304,785) Urology outpatient attendances (new and follow-up) in the Midlands region in the five-year period April 2018 to March 2023.

```{r}

table_data <-midlandsoutpatients |>
  filter(der_appointment_type != "N/A") |>
  group_by(der_financial_year,der_appointment_type) |>
  summarise(total = sum(numattends)) |>
  ungroup()

table_data|> 
  pivot_wider(names_from = der_appointment_type, values_from = total, id_cols = der_financial_year) |>
gt(rowname_col = "der_financial_year")|>
  cols_move(FUp, New) |>
  tab_header(
    title = 'Urology Outpatient Attendances - Midlands'
  ) |>
  tab_source_note(source_note = "Data: SUS")
    
```

The Urology outpatient data rarely includes diagnosis codes (only 4% of attendances). Procedures that take place in outpatients are also not consistently recorded. Consequently it's not possible to give a robust picture of activity in relation to these conditions. In the map and table below, it can be seen that over a five year peiod (April 2018 to March 2023), the ICB with the largest volume of Urology outpatient new attendances is Lincolnshire (71E) ICB with 99.61 per 1000 list size population and Leicester City (04C) has the lowest with 30.09 per 1000 list size population.

##### Urology New Outpatient Attendances per 1000 List Size

```{r}
subicblistsize <- ICBTOBPH |>
  group_by(sub_icb_location_code) |>
  summarise(sub_icb_list_size=sum(list_size_bph_total))|>
  filter(sub_icb_location_code %in% midlands_sub_icbs) |>
  ungroup()

midlandsoutpatientssubicb <-
  midlandsoutpatients |>
  filter(der_appointment_type != "N/A") |>
  group_by(sub_icb_location_code,der_appointment_type) |>
  summarise(totalProcs = sum(der_number_procedure),
            totalattends = sum(numattends)) |>
  ungroup()

midlandsoutpatientssubicb <- midlandsoutpatientssubicb |>
  left_join(subicblistsize) |>
  mutate(attendperthousand = (totalattends/sub_icb_list_size)*1000,
         procsperthousand = (totalProcs/sub_icb_list_size)*1000)


midlands_new_op_attends <- midlandsoutpatientssubicb |>
  filter(der_appointment_type == "New") |>
  left_join(subicbshape)
  
  
  tm_shape(st_sf(midlands_new_op_attends)) +
  tm_fill("attendperthousand",title="Attends per 1000",style = "pretty", palette = "-viridis") +
  tm_borders("grey25", alpha=.5) +
  tm_basemap(server = c('OpenStreetMap'))
```

```{r}
subicblistsize <- ICBTOBPH |>
  group_by(sub_icb_location_code) |>
  summarise(sub_icb_list_size=sum(list_size_bph_total))|>
  filter(sub_icb_location_code %in% midlands_sub_icbs) |>
  ungroup()

midlands_sub_icb_names <- ICBTOBPH |>
  select(sub_icb_location_code, new_sub_icb_location_name) |>
  filter(sub_icb_location_code %in% midlands_sub_icbs) |>
  filter(new_sub_icb_location_name != "NA") |>
  unique()

midlandsoutpatientssubicb <-
  midlandsoutpatients |>
  filter(der_appointment_type != "N/A") |>
  group_by(sub_icb_location_code,der_appointment_type) |>
  summarise(totalProcs = sum(der_number_procedure),
            totalattends = sum(numattends)) |>
  ungroup()

midlandsoutpatientssubicb <- midlandsoutpatientssubicb |>
  left_join(subicblistsize) |>
  left_join(midlands_sub_icb_names) |>
  mutate(attendperthousand = (totalattends/sub_icb_list_size)*1000,
         procsperthousand = (totalProcs/sub_icb_list_size)*1000)


# BARPLOT

bar_yields <- midlandsoutpatientssubicb |> 
  filter(der_appointment_type == "New") |>
  #select(-totalProcs) %>% 
  rowwise() |> 
  ungroup() |> 
  select(new_sub_icb_location_name, totalattends, attendperthousand) |> 
  mutate(
    bar = round(attendperthousand/max(attendperthousand)*100, digits = 2),
    color = col_pal(bar,"ggsci::purple_material"),
    bar_chart = bar_chart(bar, color = color),
    bar_chart = map(bar_chart, ~gt::html(as.character(.x)))
  ) |> 
  select(-bar, -color)

# BARPLOT

rule10_bar <- bar_yields |> 
  gt() |> 
  cols_width(c(bar_chart) ~ px(100),
             c(totalattends) ~ px(75),
             c(attendperthousand) ~ px(75),
  ) |> 
  cols_label(
    new_sub_icb_location_name = md("Sub-ICB"),
    totalattends = md("New Outpatient Attendances"),
    attendperthousand = md("New Outpatient Attends /1000 List Size"),
    bar_chart = ""
  ) |> 
  cols_align(
    align = "right",
    columns = 2:3
  ) |> 
  cols_align(
    align = "left",
    columns = vars(bar_chart)
  ) |> 
  fmt_number(columns = 2:2,
             decimals=0) |> 
  fmt_number(columns = 3:3,
             decimals=2) |>
  tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_column_labels(everything())
    )
  ) |>  
  tab_options(
    row_group.border.top.width = px(3),
    row_group.border.top.color = "black",
    row_group.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.top.color = "white",
    table.border.top.width = px(3),
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3),
    table_body.border.bottom.width = px(2),
    table_body.border.bottom.color = "black",
    column_labels.border.bottom.color = "black",
    column_labels.border.bottom.width = px(3)
  )
# %>% 
#   tab_footnote(footnote = "Source: SUS April 2018 - March 2023",
#                locations = cells_column_labels(
#                  columns =2:4
#                )) 

rule10_bar
```

### Prescribing

BPH

```{r}
bph_activity_icb_drugs_mid |>
  gt() |> 
  cols_hide(activityneedratio) |>
  cols_label(
    sub_icb_location_code = md("Sub-ICB"),
    bphprev = md("Prevalence"),
    list_size_bph = md("List Size"),
    num_bph_spells_sum = md("Spells"),
    items_total = md("Items"),
    presc_anr = md("Activity to Need Ratio")
  ) |>
    tab_header(
    title = 'BPH Prescription Items Activity To Need Ratio by Sub_ICB'
  ) |>
  tab_source_note(source_note = "Data: openprescribing.net")
```

OAB

```{r}
oab_activity_icb_drugs_mid |>
  gt() |> 
  cols_hide(activityneedratio) |>
  cols_label(
    sub_icb_location_code = md("Sub-ICB"),
    oabprev = md("Prevalence"),
    list_size_oab = md("List Size"),
    num_oab_spells_sum = md("Spells"),
    items_total = md("Items"),
    presc_anr = md("Activity to Need Ratio")
  ) |>
    tab_header(
    title = 'OAB Prescription Items Activity To Need Ratio by Sub_ICB'
  ) |>
  tab_source_note(source_note = "Data: openprescribing.net")
```

## Conclusion

(The medications are not particularly effective and when a patient is on multiple other medications its not necessarily a good idea to go on these drugs. Consultants say they want a patient on it but GP's then discuss with patient and when they learn of side effects etc then can decide not to (esp if elderly). Good treatment options are unfortunately limited.) - Matt

## Appendix

We will identify all relevant procedure codes and medications for inclusion and confirm these with the project's clinical contacts

### Benign Prosta Hyperplasia:

Source of data SUS APCS

Treatment Function Code: 101

Admission date: 1/4/2018 and 31/3/2023

Diagnosis codes: N40 or N32 in any diagnosis position

Procedure codes: M36, M61, M62, M65, M67, M68, M71

Elective and Day Case

Medications:

Tamsulosin hydrochloride              0704010U0         (7.4.1 drugs for urinary retention)

Tamsulosin and dutasteride hydrochloride (0704010V0)   (7.4.1 drugs for urinary retention)

Solifenacin/tamsulosin (0704010W0) (7.4.1 drugs for urinary retention) and (0704020AF) (7.4.2 Drugs for urinary frequency enuresis and incontinence) -- remove 742?

Finasteride (0604020C0) (excludes baldness as in different chapter, this one is 6.4.2 Male sex hormones and antagonists)

Alfuzosin hydrochloride (0704010A0) (7.4.1 drugs for urinary retention)

Table by GP Practice of Activity to Need Ratios for BPH

```{r}
bph_gp_mid <- ICBTOBPH |>
  filter(icb22 %in% midlands_icbs) |>
  select(org_code, practice_name, bphprevnumtotal,list_size_bph_total, num_bph_spells,new_sub_icb_location_name, icb22nm) |>
  mutate(anr = num_bph_spells/bphprevnumtotal) |>
  gt(rowname_col = "icb22nm","new_sub_icb_location_name") |>
  
    cols_label(
    org_code = md("GP Pracice Code"),
    practice_name = md("GP Practice Name"),
    bphprevnumtotal = md("Prevalence"),
    list_size_bph_total = md("List Size"),
    num_bph_spells = md("Spells"),
    anr = md("Activity To Need Ratio")
  ) |> 
    fmt_number(columns = 3:5,
             decimals=0) |> 
  tab_header(
    title = 'BPH Activity To Need Ratio by GP practice'
  ) |>
  tab_source_note(source_note = "Data: SUS & Patients Registered at a GP Practice (NHS Digital)")

bph_gp_mid

# Need to flag / remove practices with >20% variation
```

Table by GP Practice of Activity to Need Ratios for Overactive Bladder

```{r}
oab_gp_mid <- ICBTOOAB |>
  filter(icb22 %in% midlands_icbs) |>
  select(org_code, practice_name, oabprevnumtotal,list_size_oab_total, num_oab_spells,new_sub_icb_location_name, icb22nm) |>
  mutate(anr = num_oab_spells/oabprevnumtotal) |>
  gt(rowname_col = "icb22nm", "new_sub_icb_location_name") |>
    cols_label(
    org_code = md("GP Pracice Code"),
    practice_name = md("GP Practice Name"),
    oabprevnumtotal = md("Prevalence"),
    list_size_oab_total = md("List Size"),
    num_oab_spells = md("Spells"),
    anr = md("Activity To Need Ratio")
  ) |> 
      fmt_number(columns = 3:5,
             decimals=0) |> 
  tab_header(
    title = 'OAB Activity To Need Ratio by GP practice'
  ) |>
  tab_source_note(source_note = "Data: SUS & Patients Registered at a GP Practice (NHS Digital)")

oab_gp_mid

# Need to flag / remove practices with >20% variation
```

### Overactive Bladder:

ICD10 Code: N394 OPCS codes: M19, M24, M34 with X851, M36, M85, M86, A701 with Z112

Medications:

Duloxetine hydrochloride (0704020AA) (other codes are not incontinence)

Solifenacin (0704020AB)  (742)

Solifenacin/tamsulosin (0704010W0) (741) EXCLUDE?

Oxybutynin (0704020J0) (742)

Oxybutynin hydrochloride (0704040G0) (744 -- bladder instillations and urological surgery)

Tolterodine (0704020N0) (742)

Fesoterodine fumarate (0704020AD) (742)

Darifenacin hydrobromide (0704020AC) (742)

Mirabegron (0704020AE) (742)

Desmopressin acetate (0605020E0) (6.5.2 Posterior pituitary hormones and antagonists)

### Post Prostatectomy Incontinence:

ICD10 Code:

OPCS codes: M61, M65, M662, M683

Medications: Oxybutynin, Tolterodine, Solifenacin, Mirabegron, Duloxetine
